---
title: "Visual Tutorial"
author: "Jennifer Brosnahan"
date: "8/7/2020"
output:
  html_document: 
    keep_md: yes
---

```{r}
library(knitr)
library(ggplot2)
library(ggthemes)
library(scales)
library(openxlsx)
```

# Import data
```{r}
attrition <- read.csv(file.path('C:/Users/jlbro/OneDrive/R Studio projects/ggplot attrition 1', 'attrition.csv'), stringsAsFactors = TRUE)
```

# Check first 5 rows
```{r}
head(attrition)
```

# Check structure
```{r}
str(attrition)
```

### Visual 1
### 1 Essential layers

```{r}
ggplot(attrition, aes(x = JobRole, y = MonthlyIncome, fill = Attrition)) +
  geom_bar(stat = 'summary', fun = mean)
```

### 2 Coordinates and Position Adjustments

```{r}
ggplot(attrition, aes(x = JobRole, y = MonthlyIncome, fill=Attrition)) +
  geom_bar(stat = 'summary', fun = mean, position = 'dodge') +  #Unstack bars using position = ‘dodge’
  coord_flip()  #Flip x and y axis
```

### 3 Reorder Job Role by highest to lowest Monthly Income

```{r}
ggplot(attrition, aes(x = reorder(JobRole, MonthlyIncome), y = MonthlyIncome, fill = Attrition)) +
  geom_bar(stat = 'summary', fun = mean, position = 'dodge') +
  coord_flip()
```

### 4 Change colors

```{r}
ggplot(attrition, aes(x = reorder(JobRole, MonthlyIncome), y = MonthlyIncome, fill = Attrition)) +
  geom_bar(stat = 'summary', fun = mean, width = .8, position = 'dodge') + 
  coord_flip() +
  scale_fill_manual(values = c('#96adbd', '#425e72')) 
```

### 5 Add labels

```{r}
ggplot(attrition, aes(x = reorder(JobRole, MonthlyIncome), y = MonthlyIncome, fill = Attrition)) +
  geom_bar(stat = 'summary', fun = mean, width = .8, position = 'dodge') +
  coord_flip() +
  scale_fill_manual(values = c('#96adbd', '#425e72')) +
  xlab(' ') +  #Make x label invisible, notice the space between parentheses
  ylab('Monthly Income in USD') +  #Add y label
  ggtitle('Employee Attrition by Job Role & Income')  #Add title
```

### 6 Add theme

```{r}
ggplot(attrition, aes(x = reorder(JobRole, MonthlyIncome), y = MonthlyIncome, fill = Attrition)) +
  geom_bar(stat = 'summary', fun = mean, width = .8, position = 'dodge') +
  coord_flip() +
  scale_fill_manual(values = c('#96adbd', '#425e72')) +
  xlab(' ') +
  ylab('Monthly Income in USD') + 
  ggtitle('Employee Attrition by Job Role & Income') +
  theme_clean()  #Adding a theme
```

### 7 Remove outlines and minimizing aspect ratio

```{r}
ggplot(attrition, aes(x = reorder(JobRole, MonthlyIncome), y = MonthlyIncome, fill = Attrition)) +
  geom_bar(stat = 'summary', fun = mean, width = .8, position = 'dodge') +
  coord_flip() +
  scale_fill_manual(values = c('#96adbd', '#425e72')) +
  xlab('') +
  ylab('Monthly Income in USD') + 
  ggtitle('Employee Attrition by Job Role & Income') +
  theme_clean() +
  theme(aspect.ratio = .65,
        plot.background = element_rect(color = 'white'),
        legend.background = element_rect(color = 'white')) #Change height and 'remove' outlines
```

#################################################################################################

### Visual 2

### 1 Essential Layers

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE)  #Note that se = FALSE removes the confidence shading
```

### 2 Faceting to add subplots to the canvas

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~.) 
```

### 3 Add labels to facet subplots

```{r, message=FALSE}
#CreateWorkLifeBalance labels
wlb.labs <- c('1' = 'Bad Balance',
              '2' = 'Good Balance',
              '3' = 'Better Balance',
              '4' = 'Best Balance')
#Add to facet_wrap()
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~., labeller = labeller(WorkLifeBalance = wlb.labs)) 
```

### 4 Labels and Title

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~., 
             labeller = labeller(WorkLifeBalance = wlb.labs)) +
  xlab('Monthly Income') +
  ylab('Years Since Last Promotion') + 
  ggtitle('Attrition by Work-life Balance & Years Since Last Promotion')
```

### 5 Add space between labels and tick markers

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~., 
             labeller = labeller(WorkLifeBalance = wlb.labs)) +
  xlab('\nMonthly Income') +
  ylab('Years Since Last Promotion\n') + 
  ggtitle('Attrition by Work-life Balance and Promotion')
```

### 6 Theme

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~., 
             labeller = labeller(WorkLifeBalance = wlb.labs)) +
  xlab('\nMonthly Income') +
  ylab('Years Since Last Promotion\n') + 
  ggtitle('Attrition, Work-life Balance, and Years Last Promoted') +
  theme_fivethirtyeight()
```

### 7 Override theme default to bring x and y labels back

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~., 
             labeller = labeller(WorkLifeBalance = wlb.labs)) +
  xlab('\nMonthly Income') +
  ylab('Years Since Last Promotion\n') + 
  ggtitle('Attrition, Work-life Balance, and Years Last Promoted') +
  theme_fivethirtyeight() +
  theme(axis.title = element_text())  #Bringing back our x and y labels
```

### 8 Add space and change legend location

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~., 
             labeller = labeller(WorkLifeBalance = wlb.labs)) +
  xlab('\nMonthly Income') +
  ylab('Years Since Last Promotion\n') + 
  ggtitle('Attrition by Work-life Balance and Promotion') +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        legend.position = 'top',
        legend.justification = 'left',
        panel.spacing = unit(1.5, 'lines'))
```

### 9 Change line color

```{r, message=FALSE}
ggplot(attrition, aes(x = MonthlyIncome, y = YearsSinceLastPromotion, color=Attrition)) +
  geom_smooth(se = FALSE) +
  facet_wrap(WorkLifeBalance~., 
             labeller = labeller(WorkLifeBalance = wlb.labs)) +
  xlab('\nMonthly Income') +
  ylab('Years Since Last Promotion\n') + 
  ggtitle('Attrition by Work-life Balance and Promotion') +
  theme_fivethirtyeight() +
  theme(axis.title = element_text(),
        legend.position = 'top',
        legend.justification = 'left',
        panel.spacing = unit(1.5, 'lines')) +
  scale_color_manual(values = c('#999999','#ffb500'))  #Change line color
```

